Theory about Normal Basis on $F(2^k)$

### 2017-03-03 dissertation/submission/normal.tex
after submission: revision 2

\chapter{Functional Verification of Sequential Normal Basis Multipliers}
\label{ch:normal}
With the size of sequential arithmetic circuits increasing rapidly, 
conventional methods become inefficient on verifying the correctness of these circuits.
In Chapter \ref{ch:reacha} we proposed a FSM traversal approach at word-level 
based on abstractions. Since the sequential arithmetic circuits can be modeled 
as FSMs, it is possible to apply our word-level traversal algorithm to solving 
the functional verification problem on these circuits.
From the work of Lv \cite{lv:phd},
we learn that arithmetic circuits in Galois field (GF) can be
represented using an ideal of polynomials, and the 
ideal generators themselves form a GB under a specialized term order
obtained by a reverse topological traversal of the circuit.
Furthermore, according to the work of Pruss {\it et al.}
\cite{pruss:tcad15}, we can also abstract the word-level signature of an arithmetic 
component working in GF. Thus, we consider exploiting the concepts of \cite{lv:phd} and \cite{pruss:tcad15}
to efficiently perform abstraction of the state-space to specific sequential GF circuits.
In each time-frame, we can use the techniques 
from \cite{pruss:tcad15} to abstract the word-level
signature of the combinational logic, which corresponds
to the transition function in our traversal algorithm.
We show how this approach can be applied to verify sequential GF multipliers
which we can apply our traversal algorithm to actually 
verify its functional correctness.

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/Moore.eps}
\caption{A typical Moore machine and its state transition graph.}
\label{fig:Moore}}
\end{figure}

In practice,
some arithmetic components are designed in sequential circuits similar to the structure in 
Figure \ref{fig:Moore}(a). Initially the operands are loaded into the registers, 
then the closed circuit executes without taking any additional information from outside,
and stores the results in registers after $k$ clock cycles. Its behavior can be described using the 
STG in Figure \ref{fig:Moore}(b): state $R$ denotes the bits stored in registers. Concretely, 
$R_{init}$ is the initial state, $R_1$ to $R_{k-1}$ are intermediate results stored as NS variables 
of current time-frame and PS variables for the next time-frame, and $R_k$ (or $R_{final}$) 
is the final result given by arithmetic circuits (and equals to the
arithmetic function when circuit is working correctly).
This kind of design results in 
reusing a smaller combinational logic component such that the area cost is greatly optimized.
An example of such designs is the sequential GF arithmetic multipliers described in Chapter \ref{ch:prelim_GF}.
However, it also brings difficulties in verifying the the circuit functions.

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/convention.eps}
\caption{Conventional verification techniques based on bit-level unrolling and equivalence checking.}
\label{fig:convention}}
\end{figure}

Conventional methods to such a sequential circuit may consist of unrolling the circuit for 
$k$ time-frames, and performing an equivalence checking between the unrolled machine and
the specification function. However, the number of gates will grow fast when unrolling
on bit-level. Meanwhile the structural similarity based equivalence checking techniques 
will fail when the sequential circuit is highly customized and optimized from the naive specification 
function. As a result, conventional techniques are grossly inefficient for large circuits.
Therefore, a new method based on our proposed word-level FSM traversal technique is worthy of exploration.

\section{Formal Verification of Normal Basis Multipliers using Gr\"obner Bases}
The gate-level design of a NB multiplier can be generated using approaches introduced in 
Section \ref{sec:nbdesign}. The gate-level netlist is ready to be verified using 
an approach similar to that in Chapter \ref{ch:reacha}. 
First we introduce the sketch of our approach using abstraction term order (ATO)
mentioned previously in Section \ref{sec:abstraction}, then refine our approach using 
the concept of ``RATO", which is previously used in Section \ref{sec:improve}.
% The following part borrows contents
% from my own conference paper \cite{myDATE}.
\subsection{Implicit Unrolling based on Abstraction with ATO}
The concept of abstraction was discussed in Section \ref{sec:abstraction}. If we use an elimination term order with
$intermediate~variables~>~R~>~A,B$ for the circuit in Figure \ref{fig:comb_RAB} to compute Gr\"obner basis, 
the function of the combinational logic component can be abstracted as 
$$R = \Func(A,B)$$
To verify the functional correctness of a combinational NB multiplier ({\it e.g.} Mastrovito multiplier
or Montgomery multiplier), the function given by abstraction will be computed as
$$R=A\cdot B$$
While in the sequential case, the function of combinational logic only fulfills a part of the multiplication.
For example, in the RH-SMPO introduced in Section \ref{sec:nbdesign}, 
the combinational logic actually implements  $F_{k-1}$ in Equation \ref{eqn:aux},
while computing all of $F_i$ requires $k$-cycle unrolling of the circuit.
Nevertheless, the abstraction still provides a word-level 
representation which can be more efficient for unrolling than bit-level expressions.
In other words, with the assistance of abstraction, we can execute implicit unrolling instead of 
explicit unrolling and avoid bit-blasting problem.

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/comb_RAB.pdf}
\caption{Architecture of a combinational GF multiplier.}
\label{fig:comb_RAB}}
\end{figure}

@@

For 2-input sequential NB multipliers, abstraction is utilized to implement Algorithm \ref{alg:modified}.
Notations in Figure \ref{fig:sequential} as well as in next example comply with 
the notations in this algorithm.

\begin{figure}[bp]
\centering{
\includegraphics[width=5in]{newfig/gf_seq_model.eps}
\caption{A typical normal basis GF sequential circuit model. $A =
  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
  $A', B', R'$ denote next-state inputs.}
\label{fig:sequential}}
\end{figure}

@@

Each gate in the combinational logic is represented by a Boolean
polynomial. To 
this set of Boolean polynomials, we append the polynomials that define
the word-level to bit-level relations for PS and NS variables ($A =
\sum_{i=0}^{k-1} a_i \beta^{2^i}$). We denote this set of polynomials
as ideal $J_{ckt} = \langle 
f_1, \dots, f_s \rangle \subset \F_2[x_1, \dots, x_d, R, R', A, A', B,
  B']$, where $x_1, \dots, x_d$ denote the bit-level (Boolean) variables
  of the circuit. The ideal of vanishing polynomials 
$J_0 = \langle x_i^2-x_i,R_j^{2^k}-R_j\rangle,~\forall \text{ bit-level variables }x_i\text{ and word-level variables }R_j$, is also included. 
Then the implicit FSM unrolling problem is setup for abstraction.

The configurations of the flip-flops are the states of the
machine. Since the set of states is a finite set of points, we
can consider it as the variety of an ideal related to the circuit.
Moreover, since we are interested in
the function encoded by the state variables (over $k$-time
frames), we can project this variety on the word-level state
variables, starting from the initial state $A_{init}, B_{init}$.
Projection of varieties (geometry) corresponds to elimination ideals
(algebra), and can be analyzed via \Grobner bases as shown in Section \ref{sec:abstraction}. Therefore, we
employ a \Grobner basis computation with ATO: we use a {\it lex term
  order} with {\it bit-level variables} 
$>$ {\it word-level NS outputs} $>$ {\it word-level PS inputs}. This
allows to eliminate all the bit-level variables 
%(corresponding to the combinational logic and the state variables),
%so as to 
and derives a representation only in terms of words. 
Consequently, $k$-successive \Grobner basis computations implicitly
unroll the machine, and provide a word-level algebraic $k$-cycle
abstraction for $R'$ as $R' = \Func(A_{init}, B_{init})$.

@@

In the practice of sequential GF multiplier verification, the combinational 
logic actually implements function not only related to current operands $A$ and $B$,
but also involved with PS variable ({\it i.e.} temporary product) $R$. Which can be 
obtained using the abstraction:
$$R' = \Func(A,B,R)$$
Using ATO, if we put $R$ ahead of $R',A,B$ in term ordering, $R$ is thus eliminated
and the result in single iteration will be $R' = \Func(A,B)$. NS operands 
$A'$ and $B'$ are right-cyclic shift of $A$ and $B$, which can be directly written.
If initial values $A_{init}, B_{init}$ are treated as parameters, the NS ideal 
contains polynomials with $A', B'$ and $R' = \Func(A_{init},B_{init})$. This 
is also shown in the following example.

\begin{Example}[Functional verification of 5-bit RH-SMPO]
\label{ex:RHSMPO}
Figure \ref{fig:5bitRH} shows the detailed structure of a 5-bit RH-SMPO, with normal element $\beta = \alpha^5$. The transition function for
operands $A,B$ perform cyclic shift, while the transition function for $R$ has to be computed through the Gr\"obner basis
abstraction approach. Following ideal $J_{ckt}$ from line 5 in Algorithm \ref{alg:modified} is the ideal for 
all gates in combinational logic block and definition of word-level variables.
$$J_{ckt} = \langle f_1,f_2,\dots,f_{19}\rangle$$
\begin{eqnarray*}
 \left. 
\begin{aligned}
d_0+a_4b_4, c_1+a_0+a_4, c_2+b_0+b_4, d_1+c_1c_2, c_3+a_1a_4\\
c_4+b_1b_4, d_2+c_3c_4, e_0+d_0+d_1, e_3+d_1+d_2, e_4+d_2 \\
R_0+r_4+e_0, R_1+r_0, R_2+r_1, R_3+r_2+e_3, R_4+r_3+e_4
\end{aligned}\right\} ~~~\{f_1\dots f_{15}\}\\
 \left.
\begin{aligned}
 A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}\\
		  B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}\\
		  R'+r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18}\\
		  R+R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18}
		 \end{aligned}\right\} ~~~\{f_{16}\dots f_{19}\}
\end{eqnarray*}

@@

According to line 5 of Algorithm \ref{alg:modified}, we merge $J_{ckt}$, $J_0$ and $from^i$, then compute its
Gr\"obner basis with abstraction term order. 
Concretely, in the first iteration $from^0$ contains three generators. The first one describes PS variable $R$ --
temporary (or initial, in first iteration) product, which equals to 0 according to the mechanism of sequential 
GF multiplier. The polynomial is written as 
$$R$$
The other two polynomials describes PS variable $A,B$ -- current multiplication operands, and we write them 
in parameters $A_{init}, B_{init}$. Since this is the first iteration, they have the same form with $f_{16}$ and $f_{17}$:
\begin{align*}
& A_{init}+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18} \\
& B_{init}+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}
\end{align*}
After computing the Gr\"obner basis of $J_{ckt}+J_0+from^0$ using ATO 
$$\text{all bit-level variables} > R > \{R',A_{init},B_{init}\}$$
there is a polynomial in the form of $R'+\Func(A_{init},B_{init})$,
which should be included in $to^{i+1}$. $to^{i+1}$ also excludes next state variable $A'$ and $B'$, instead we 
redefine $A_{init}$ and $B_{init}$ using next state bit-level variables $\{a_i', b_j'\}$. Next state Bit-level variables
$a_i' = a_{i-1\pmod k}, b_j' = b_{j-1\pmod k}$ according to definition of cyclic shift.

@@

\subsection{Overcome Computational Complexity using RATO}
We implemented Algorithm \ref{alg:modified} in \textsc{Singular}, with similar setup and environments as
with the experiments in Section \ref{sec:exp_reacha}. The result of verifying Agnew's SMPO is shown 
in Table \ref{tab:slimgb}.

@@

It indicates that our approach based on ATO cannot verify sequential GF multiplier with size larger than 11 bits operands.
Similar to our improvements in Section \ref{sec:improve}, RATO \cite{TimDAC} can also be applied to 
accelerate the GB computation here. More specifically, we circumvent the GB computation 
of Line 5 in Algorithm \ref{alg:modified} by turning the Buchberger's algorithm 
into a single-step of multivariate polynomial division.

\begin{itemize}
\item First, a set of polynomials is constructed from each logic gate in the circuit, and they generate 
ideal $J_{gates}$. Similarly, ideal $J_{WB}$ is generated by polynomials representing the word-to-bit correspondences.
They are merged as an ideal describing the combinational logic of the circuit: $J_{ckt} = J_{gates} + J_{WB}$.
\item Second, we impose RATO on ideal $J_{ckt}$: all bit-level variables are sorted using a reverse topological order
derived from the circuit structure, followed by word-level NS outputs ({\it e.g.} $R'$ in Figure \ref{fig:sequential}), 
then word-level PS inputs ({\it e.g.} $A,B,R$ in Figure \ref{fig:sequential}).
\item When computing the Gr\"obner basis of $J_{ckt}+J_0$ (adding vanishing polynomials), because of 
RATO, there exists only one pair of polynomials $(f_w,f_g)$ such that $LCM(lm(f_w),lm(f_g)) \neq lm(f_w)\cdot lm(f_g)$.
\item Then we only need to compute the $Spoly$ for $(f_w,f_g)$, and reduce the $Spoly$ by $J_{ckt}+J_0$.
\end{itemize}

@@

\begin{Example}
\label{ex:newRATO}
From the circuit topological structure in Figure \ref{fig:5bitRH}, the term order under RATO is:
\begin{align}
&\{r_0',r_1',r_2',r_3',r_4'\}>\{r_0,r_1,r_2,r_3,r_4\}\nonumber\\
&>\{e_0,e_3,e_4\},\{d_0,d_1,d_2\},\{c_1,c_2,c_3,c_4\}\nonumber\\
&>\{a_0,a_1,a_2,a_3,a_4,b_0,b_1,b_2,b_3,b_4\}>R'>R>\{A,B\}\nonumber
\end{align}
The variables in braces are at the same topological level, they are arranged in LEX order. 
We search among all generators of $J_{ckt}$ from Example \ref{ex:RHSMPO} using RATO, 
and find a pair of polynomials whose leading 
monomials are not relatively prime:
$(f_w,f_g), f_w = r_0'+r_4+e_0, f_g =r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18} + R'$.
We calculate $Spoly$ can reduce it by $J_{ckt}+J_0$:
\begin{align}
&S\text{-}poly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}\nonumber\\
&(\alpha^3+\alpha^2+\alpha) r_1+(\alpha^4+\alpha^3+\alpha^2) r_2+(\alpha^2+\alpha) r_3+(\alpha) r_4\nonumber\\
&+(\alpha^3+\alpha^2) a_1 b_1+(\alpha^4+\alpha^3+\alpha^2+\alpha) a_1 b_2+(\alpha^2+\alpha) a_1 b_3\nonumber\\
&+(\alpha^2+1) a_1 b_4+(\alpha^4+1) a_1 B+(\alpha^4+\alpha) a_2 b_1+(\alpha^4+\alpha^3+\alpha) a_2 b_2\nonumber\\
&+(\alpha^3+1) a_2 b_3+(\alpha^3+\alpha^2+1) a_2 b_4+(\alpha^3+\alpha^2) a_2 B+(\alpha^2+\alpha) a_3 b_1\nonumber\\
&+(\alpha^3+1) a_3 b_2+(\alpha+1) a_3 b_3+(\alpha^4+\alpha^2+\alpha) a_3 b_4\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) a_3 B+(\alpha^3+1) a_4 b_1+a_4 b_2+(\alpha^4+\alpha^2+\alpha) a_4 b_3\nonumber\\
&+(\alpha^4+\alpha^3+1) a_4 b_4+(\alpha^2+\alpha) a_4 B+(\alpha^4+1) b_1 A+(\alpha^3+\alpha^2) b_2 A\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) b_3 A+(\alpha^2+\alpha) b_4 A+(\alpha^4+\alpha^2+\alpha+1) R'+R+A B\nonumber
\end{align}
\end{Example}
The above example indicates that RATO based abstraction on 5-bit RH-SMPO will result in a remainder containing both
bit-level variables and word-level variables. Moreover, the number of variables is still large such that
Gr\"obner basis computation will be inefficient.

Since the remainder from \emph{Spoly} reduction contains some bit-level variables, our objective is to compute
a polynomial that contains only word-level variables (such as $R'+\mathcal{F}(A,B)$). One possible solution to
this problem is to replace the bit-level variables by equivalent polynomials that only contain word-level
variables, {\it e.g.} $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
approach is introduced to compute corresponding $\mathcal{G}(A),\mathcal{H}(R)$ efficiently.

@@

The \textsc{Singular} tool can read scripts written in its own format similar to ANSI-C. 
The input file format for our C++ tool is also designed to be compatible with \textsc{Singular} 
ring and polynomial definitions. For SMPO experiments, the main 
loop of our script file performs the same function as Algorithm \ref{alg:modified} describes, while the Gr\"obner basis
computation in main loop is divided into 4 different functional parts:

1) Pre-processing:

>>>>

\chapter{Functional Verification of Sequential Normal Basis\protect\\ Multipliers}
\label{ch:normal}
With the size of sequential arithmetic circuits increasing rapidly, 
conventional methods become inefficient for verifying the correctness of these circuits.
In Chapter \ref{ch:reacha} we proposed a FSM traversal approach at word level 
based on abstractions. Since the sequential arithmetic circuits can be modeled 
as FSMs, it is possible to apply our word-level traversal algorithm to solving 
the functional verification problem on these circuits.
From the work of Lv \cite{lv:phd},
we learn that arithmetic circuits in Galois field (GF) can be
represented using an ideal of polynomials, and the 
ideal generators themselves form a GB under a specialized term order
obtained by a reverse topological traversal of the circuit.
Furthermore, according to the work of Pruss {\it et al.}
\cite{pruss:tcad15}, we can also abstract the word-level signature of an arithmetic 
component working in GF. Thus, we consider exploiting the concepts of \cite{lv:phd} and \cite{pruss:tcad15}
to efficiently perform abstraction of the state-space to specific sequential GF circuits.
In each time-frame, we can use the techniques 
from \cite{pruss:tcad15} to abstract the word-level
signature of the combinational logic, which corresponds
to the transition function in our traversal algorithm.
We show how this approach can be applied to verify sequential GF multipliers
which we can apply to our traversal algorithm to actually 
verify its functional correctness.

@@

In practice,
some arithmetic components are designed in sequential circuits similar to the structure in 
Figure \ref{fig:Moore}(a). Initially the operands are loaded into the registers, 
then the closed circuit executes without taking any additional information from outside,
and stores the results in registers after $k$ clock cycles. Its behavior can be described using the 
STG in Figure \ref{fig:Moore}(b): state $R$ denotes the bits stored in registers. Concretely, 
$R_{init}$ is the initial state, $R_1$ to $R_{k-1}$ are intermediate results stored as NS variables 
of current time-frame and PS variables for the next time-frame, and $R_k$ (or $R_{final}$) 
is the final result given by arithmetic circuits (and equals to the
arithmetic function when circuit is working correctly).
This kind of design results in 
reusing a smaller combinational logic component such that the area cost is greatly optimized.
An example of such designs is the sequential GF arithmetic multipliers described in Chapter \ref{ch:prelim_GF}.
However, it also brings difficulties in verifying the circuit functions.

\begin{figure}[tbp]
\centering{
\includegraphics[width=\textwidth]{newfig/Moore.eps}
\caption{A typical Moore machine and its state transition graph.}
\label{fig:Moore}}
\end{figure}

Conventional methods to such a sequential circuit may consist of unrolling the circuit for 
$k$ time-frames, and performing an equivalence checking between the unrolled machine and
the specification function, as shown in Figure \ref{fig:convention}. However, the number of gates will grow fast when unrolling
at bit level. Meanwhile the structural similarity-based equivalence checking techniques 
will fail when the sequential circuit is highly customized and optimized from the naive specification 
function. As a result, conventional techniques are grossly inefficient for large circuits.
Therefore, a new method based on our proposed word-level FSM traversal technique is worthy of exploration.

\begin{figure}[tbp]
\centering{
\includegraphics[width=\textwidth]{newfig/convention.eps}
\caption{Conventional verification techniques based on bit-level unrolling and equivalence checking.}
\label{fig:convention}}
\end{figure}

\section{Formal Verification of Normal Basis Multipliers Using Gr\"obner Bases}
The gate-level design of a NB multiplier can be generated using approaches introduced in 
Section \ref{sec:nbdesign}. The gate-level netlist is ready to be verified using 
an approach similar to that in Chapter \ref{ch:reacha}. 
First we introduce the sketch of our approach using abstraction term order (ATO),
mentioned previously in Section \ref{sec:abstraction}, then refine our approach using 
the concept of ``RATO", which was previously used in Section \ref{sec:improve}.
% The following part borrows contents
% from my own conference paper \cite{myDATE}.
\subsection{Implicit Unrolling Based on Abstraction with ATO}
The concept of abstraction was discussed in Section \ref{sec:abstraction}. If we use an elimination term order with
$intermediate~variables~>~R~>~A,B$ for the circuit in Figure \ref{fig:comb_RAB} to compute Gr\"obner basis, 
the function of the combinational logic component can be abstracted as 
$$R = \Func(A,B)$$
To verify the functional correctness of a combinational NB multiplier (e.g., Mastrovito multiplier
or Montgomery multiplier), the function given by abstraction will be computed as
$$R=A\cdot B$$
While in the sequential case, the function of combinational logic only fulfills a part of the multiplication.
For example, in the RH-SMPO introduced in Section \ref{sec:nbdesign}, 
the combinational logic actually implements  $F_{k-1}$ in Equation \ref{eqn:aux},
while computing all of $F_i$ requires $k$-cycle unrolling of the circuit.
Nevertheless, the abstraction still provides a word-level 
representation which can be more efficient for unrolling than bit-level expressions.
In other words, with the assistance of abstraction, we can execute implicit unrolling instead of 
explicit unrolling and avoid a bit-blasting problem.

\begin{figure}[tbp]
\centering{
\includegraphics[width=\textwidth]{newfig/comb_RAB.pdf}
\caption{Architecture of a combinational GF multiplier.}
\label{fig:comb_RAB}}
\end{figure}

@@

For 2-input sequential NB multipliers, abstraction is utilized to implement Algorithm \ref{alg:modified}.
Notations in Figure \ref{fig:sequential} as well as in the next example comply with 
the notations in this algorithm.

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=4.5in]{newfig/gf_seq_model.eps}
\caption{A typical normal basis GF sequential circuit model. $A =
  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
  $A', B', R'$ denote next-state inputs.}
\label{fig:sequential}}
\end{figure}

Each gate in the combinational logic is represented by a Boolean
polynomial. To 
this set of Boolean polynomials, we append the polynomials that define
the word-level to bit-level relations for PS and NS variables ($A =
\sum_{i=0}^{k-1} a_i \beta^{2^i}$). We denote this set of polynomials
as ideal $J_{ckt} = \langle 
f_1, \dots, f_s \rangle \subset \F_2[x_1, \dots, x_d, R, R', A, A', B,
  B']$, where $x_1, \dots, x_d$ denote the bit-level (Boolean) variables
  of the circuit. The ideal of vanishing polynomials 
$J_0 = \langle x_i^2-x_i,R_j^{2^k}-R_j\rangle,~\forall \text{ bit-level variables }x_i\text{, and word-level variables }R_j$, is also included. 
Then the implicit FSM unrolling problem is set up for abstraction.

The configurations of the flip-flops are the states of the
machine. Since the set of states is a finite set of points, we
can consider it as the variety of an ideal related to the circuit.
Moreover, since we are interested in
the function encoded by the state variables (over $k$-time
frames), we can project this variety on the word-level state
variables, starting from the initial state $A_{init}, B_{init}$.
Projection of varieties (geometry) corresponds to elimination ideals
(algebra), and can be analyzed via \Grobner bases as shown in Section \ref{sec:abstraction}. Therefore, we
employ a \Grobner basis computation with ATO: we use a {\it lex term
  order} with {\it bit-level variables} 
$>$ {\it word-level NS outputs} $>$ {\it word-level PS inputs}. This
allows us to eliminate all the bit-level variables 
%(corresponding to the combinational logic and the state variables),
%so as to 
and derives a representation only in terms of words. 
Consequently, $k$-successive \Grobner basis computations implicitly
unroll the machine, and provide a word-level algebraic $k$-cycle
abstraction for $R'$ as $R' = \Func(A_{init}, B_{init})$.

@@

In the practice of sequential GF multiplier verification, the combinational 
logic actually implements function not only related to current operands $A$ and $B$,
but also involved with PS variable (i.e., temporary product) $R$, which can be 
obtained using the abstraction:
$$R' = \Func(A,B,R)$$
Using ATO, if we put $R$ ahead of $R',A,B$ in term ordering, $R$ is thus eliminated
and the result in single iteration will be $R' = \Func(A,B)$. NS operands 
$A'$ and $B'$ are right-cyclic shift of $A$ and $B$, which can be directly written.
If initial values $A_{init}, B_{init}$ are treated as parameters, the NS ideal 
contains polynomials with $A', B'$ and $R' = \Func(A_{init},B_{init})$. This 
is also shown in the following example.

\begin{Example}[Functional verification of 5-bit RH-SMPO]
\label{ex:RHSMPO}
Figure \ref{fig:5bitRH} shows the detailed structure of a 5-bit RH-SMPO, with normal element $\beta = \alpha^5$. The transition function for
operands $A,B$ perform cyclic shift, while the transition function for $R$ has to be computed through the Gr\"obner basis
abstraction approach. Following ideal $J_{ckt}$ from line 5 in Algorithm \ref{alg:modified} is the ideal for 
all gates in a combinational logic block and definition of word-level variables.
$$J_{ckt} = \langle f_1,f_2,\dots,f_{19}\rangle$$
\begin{eqnarray*}
 \left. 
\begin{aligned}
d_0+a_4b_4, c_1+a_0+a_4, c_2+b_0+b_4, d_1+c_1c_2, c_3+a_1a_4\\
c_4+b_1b_4, d_2+c_3c_4, e_0+d_0+d_1, e_3+d_1+d_2, e_4+d_2 \\
R_0+r_4+e_0, R_1+r_0, R_2+r_1, R_3+r_2+e_3, R_4+r_3+e_4
\end{aligned}\right\} ~~~\{f_1\dots f_{15}\}\\
 \left.
\begin{aligned}
 A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}\\
		  B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}\\
		  R'+r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18}\\
		  R+R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18}
		 \end{aligned}\right\} ~~~\{f_{16}\dots f_{19}\}
\end{eqnarray*}

@@

According to line 5 of Algorithm \ref{alg:modified}, we merge $J_{ckt}$, $J_0$ and $from^i$, then compute its
Gr\"obner basis with abstraction term order. 
Concretely, in the first iteration $from^0$ contains three generators. The first one describes PS variable $R$ --
temporary (or initial, in first iteration) product, which equals to 0 according to the mechanism of sequential 
GF multiplier. The polynomial is written as 
$$R$$
The other two polynomials describe PS variable $A,B$ -- current multiplication operands, and we write them 
in parameters $A_{init}, B_{init}$. Since this is the first iteration, they have the same form as $f_{16}$ and $f_{17}$:
\begin{align*}
& A_{init}+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18} \\
& B_{init}+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}
\end{align*}
After computing the Gr\"obner basis of $J_{ckt}+J_0+from^0$ using ATO 
$$\text{all bit-level variables} > R > \{R',A_{init},B_{init}\}$$
there is a polynomial in the form of $R'+\Func(A_{init},B_{init})$,
which should be included in $to^{i+1}$. $to^{i+1}$ also excludes next state variable $A'$ and $B'$, instead we 
redefine $A_{init}$ and $B_{init}$ using next state bit-level variables $\{a_i', b_j'\}$. Next state Bit-level variables
$a_i' = a_{i-1\pmod k}, b_j' = b_{j-1\pmod k}$ according to definition of cyclic shift.

@@

\subsection{Overcome Computational Complexity Using RATO}
We implemented Algorithm \ref{alg:modified} in \textsc{Singular}, with similar setup and environments as
with the experiments in Section \ref{sec:exp_reacha}. The result of verifying Agnew's SMPO is shown 
in Table \ref{tab:slimgb}. It indicates that our approach based on ATO cannot verify sequential GF multiplier with size larger than 11 bits operands.
Similar to our improvements in Section \ref{sec:improve}, RATO \cite{TimDAC} can also be applied to 
accelerate the GB computation here. More specifically, we circumvent the GB computation 
of Line 5 in Algorithm \ref{alg:modified} by turning the Buchberger's algorithm 
into a single-step of multivariate polynomial division.

@@

\begin{itemize}
\item First, a set of polynomials is constructed from each logic gate in the circuit, and they generate 
ideal $J_{gates}$. Similarly, ideal $J_{WB}$ is generated by polynomials representing the word-to-bit correspondences.
They are merged as an ideal describing the combinational logic of the circuit: $J_{ckt} = J_{gates} + J_{WB}$.
\item Second, we impose RATO on ideal $J_{ckt}$: all bit-level variables are sorted using a reverse topological order
derived from the circuit structure, followed by word-level NS outputs (e.g., $R'$ in Figure \ref{fig:sequential}), 
then word-level PS inputs ({\it e.g.} $A,B,R$ in Figure \ref{fig:sequential}).
\item When computing the Gr\"obner basis of $J_{ckt}+J_0$ (adding vanishing polynomials), because of 
RATO, there exists only one pair of polynomials $(f_w,f_g)$ such that $LCM(lm(f_w),lm(f_g)) \neq lm(f_w)\cdot lm(f_g)$.
\item Then we only need to compute the $Spoly$ for $(f_w,f_g)$, and reduce the $Spoly$ by $J_{ckt}+J_0$.
\end{itemize}

@@

\begin{Example}
\label{ex:newRATO}
From the circuit topological structure in Figure \ref{fig:5bitRH}, the term order under RATO is:
\begin{align}
&\{r_0',r_1',r_2',r_3',r_4'\}>\{r_0,r_1,r_2,r_3,r_4\}\nonumber\\
&>\{e_0,e_3,e_4\},\{d_0,d_1,d_2\},\{c_1,c_2,c_3,c_4\}\nonumber\\
&>\{a_0,a_1,a_2,a_3,a_4,b_0,b_1,b_2,b_3,b_4\}>R'>R>\{A,B\}\nonumber
\end{align}
The variables in braces are at the same topological level and are arranged in LEX order. 
We search among all generators of $J_{ckt}$ from Example \ref{ex:RHSMPO} using RATO, 
and find a pair of polynomials whose leading 
monomials are not relatively prime:
$(f_w,f_g), f_w = r_0'+r_4+e_0, f_g =r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18} + R'$.
We calculate $Spoly$ can reduce it by $J_{ckt}+J_0$:
\begin{align}
&S\text{-}poly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}\nonumber\\
&(\alpha^3+\alpha^2+\alpha) r_1+(\alpha^4+\alpha^3+\alpha^2) r_2+(\alpha^2+\alpha) r_3+(\alpha) r_4\nonumber\\
&+(\alpha^3+\alpha^2) a_1 b_1+(\alpha^4+\alpha^3+\alpha^2+\alpha) a_1 b_2+(\alpha^2+\alpha) a_1 b_3\nonumber\\
&+(\alpha^2+1) a_1 b_4+(\alpha^4+1) a_1 B+(\alpha^4+\alpha) a_2 b_1+(\alpha^4+\alpha^3+\alpha) a_2 b_2\nonumber\\
&+(\alpha^3+1) a_2 b_3+(\alpha^3+\alpha^2+1) a_2 b_4+(\alpha^3+\alpha^2) a_2 B+(\alpha^2+\alpha) a_3 b_1\nonumber\\
&+(\alpha^3+1) a_3 b_2+(\alpha+1) a_3 b_3+(\alpha^4+\alpha^2+\alpha) a_3 b_4\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) a_3 B+(\alpha^3+1) a_4 b_1+a_4 b_2+(\alpha^4+\alpha^2+\alpha) a_4 b_3\nonumber\\
&+(\alpha^4+\alpha^3+1) a_4 b_4+(\alpha^2+\alpha) a_4 B+(\alpha^4+1) b_1 A+(\alpha^3+\alpha^2) b_2 A\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) b_3 A+(\alpha^2+\alpha) b_4 A+(\alpha^4+\alpha^2+\alpha+1) R'+R+A B\nonumber
\end{align}
\end{Example}
The above example indicates that RATO-based abstraction on 5-bit RH-SMPO will result in a remainder containing both
bit-level variables and word-level variables. Moreover, the number of variables is still large such that
Gr\"obner basis computation will be inefficient.

Since the remainder from \emph{Spoly} reduction contains some bit-level variables, our objective is to compute
a polynomial that contains only word-level variables (such as $R'+\mathcal{F}(A,B)$). One possible solution to
this problem is to replace the bit-level variables by equivalent polynomials that only contain word-level
variables, e.g. $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
approach is introduced to compute corresponding $\mathcal{G}(A),\mathcal{H}(R)$ efficiently.

@@

The \textsc{Singular} tool can read scripts written in its own format similar to ANSI-C. 
The input file format for our C++ tool is also designed to be compatible with \textsc{Singular} 
ring and polynomial definitions. For SMPO experiments, the main 
loop of our script file performs the same function as Algorithm \ref{alg:modified} describes, while the Gr\"obner basis
computation in the main loop is divided into 4 different functional parts:

1) Preprocessing:
